Bias in the Estimation of the Value of VRA over URA: A Case of Mis-speficied Geographically Weighted Regression Models

### 2023-11-15 Writing/manuscript_gwr.rmd
old paragraph back

GWR has been presented as a valuable tool for the estimation of site-specific yield response functions used to derive site-specific input rate recommendations from data generated by OFPE [@rakshit2020; @trevisan2021spatial]. @trevisan2021spatial, in particular, reported the economic value of GWR-based variable rate application (VRA) over the optimal uniform rate application (URA) in corn nitrogen management. However, the existing GWR estimations of yield response functions are using a simple linear or quadratic functional form specification. Numerous past agronomic studies recognize that crop yield responses usually follow nonlinear functional forms. For example, corn yield follows a quadratic-plateau response to nitrogen, yet a simple quadratic model was employed by @trevisan2021spatial. Ideally, the GWR model should be specified with quadratic-plateau or more flexible functional forms. However, no software package is readily usable to estimate GWR with quadratic-plateau functional form. The currently used quadratic GWR models are mostly likely mis-specified for yield response functions. Furthermore, GWR itself has faced criticism regarding the statistical reliability of its coefficient estimations [@griffith2008spatial; @harris2019simulation; @paez2011simulation; @da2016multiple]. We hypothesize that the intersection of these two issues—the model mis-specification and the inherent limitations of GWR—may have led to a substantial overestimation of the economic value of the GWR-based VRA over URA. Given the notably high economic value ($58 per hectare) of GWR-based VRA reported in @trevisan2021spatial compared to previous studies on VRA economic evaluation [@bullock1998does; @koch2004economic; @thrikawala1999economic], our suspicion stands a good chance of being plausible.

>>>>

GWR has been presented as a valuable tool for the estimation of site-specific yield response functions used to derive site-specific input rate recommendations from data generated by OFPE [@rakshit2020; @trevisan2021spatial]. @trevisan2021spatial, in particular, reported an unusually high potential gain of $58 per hectare from transitioning from the optimal uniform application rate to optimal site-specific application rates. Such a figure is notably higher than what has been suggested by previous studies on the value of variable rate application (VRA) over uniform rate application (URA) [@bullock1998does; @koch2004economic; @thrikawala1999economic]. We suspect that their reported estimate for the benefit of GWR-based VRA over URA may be significantly over-estimated for the very reasons mentioned earlier. First, their model for corn yield response to nitrogen is mostly likely mis-specified. Numerous past agronomic studies recognize that corn yield follows a quadratic-plateau response to nitrogen, yet a simple quadratic model was employed by @trevisan2021spatial. Similarly, @rakshit2020 also used quadratic model. Ideally, the GWR model should be specified with quadratic-plateau or more flexible functional forms. However, no software package is readily usable to estimate GWR with quadratic-plateau fucntional form. Second, GWR itself has faced criticism regarding the statistical reliability of its coefficient estimations [@griffith2008spatial; @harris2019simulation; @paez2011simulation; @da2016multiple]. We hypothesize that the intersection of these two issues—the model mis-specification and the inherent limitations of GWR—may have led to a substantial overestimation of the economic value of switching from URA to VRA.
